The globular cluster system of the Auriga simulations

### 2020-06-02 AurigaGCS_20200513.tex
License + changes in proof

\section{Introduction}
Globular star clusters (GC)s are old, bright, and ubiquitous. Various properties of
GC systems (GCS) show correlations with their host galaxies. GCSs are believed to
retain information about the galactic (gas) conditions at their times of formation.
Thus they could offer a unique insight into the (chemodynamical) evolution of their 
parent galaxies, if the formation and evolution of GCs- and GC systems themselves 
is adequately understood. However, despite decades of research (e.g.~\citealt{
1991ARA&A..29..543H, Harris2001, 2004Natur.427...31W, 2006ARA&A..44..193B, 
2012A&ARv..20...50G, 2014CQGra..31x4006K, 2018RSPSA.47470616F}), consensus on an 
exhaustive picture of the formation of GCs is yet to be reached. Most galaxies host
a GCS with a bimodal colour distribution that may reflect a bimodal metallicity 
distribution which is argued to arise from two distinct formation channels 
\citep{2006ARA&A..44..193B}. GCs are typically referred to as `blue' or `red' \citep[e.g.][]{
1985ApJ...293..424Z,1999AJ....118.1526G,2001AJ....121.2974L} where the former class
is more metal-poor, more radially extended, and showing little to no rotation wheras
the later is more metal-rich, radially concentrated, and rotating with the galaxy.

@@

Yet another hypothesis is that star cluster formation is triggered by major 
gas-rich (spiral) galaxy mergers \citep{1987nngp.proc...18S, 1992ApJ...384...50A},
which is naturally expected within the framework of hierarchical assembly. One
testable prediction of this scenario is the formation of young clusters in
interacting and merging galaxies, which has been observed and are found to show 
remarkable similarities with GCs in the Milky Way (MW) \citep[e.g.][]{
1995AJ....109..960W, 1996AJ....112..416H, 1999AJ....118..752Z, 1999AJ....118.1551W}.
Moreover, modelling efforts of this framework yield GC (sub)populations consistent 
with various observables \citep[e.g.][]{2010ApJ...718.1266M, 2018MNRAS.480.2343C}, 
and the recent numerical simulation of an isolated dwarf-dwarf merger executed at 
very high resolution (baryonic mass $m_b \sim$~4M\Sun; softening $\epsilon = 0.1$~pc) 
produces star clusters that could be GC progenitors \citep{
2019arXiv190509840L}.

@@

The (mass) resolution of cosmological zoom simulations has reached the mass range 
populated by GCs, and the gravitational force softening can be as low several 
parsec. A number of groups can thus incorporate formation of (globular) star 
clusters into their high-resolution hydrodynamical simulations. \citet{2016ApJ...831..204R} 
ran parsec scale simulations of the high-redshift universe 
prior to reionization (the simulations stop at $z=9$), \citet{2017ApJ...834...69L} 
implement a new subgrid model for star (cluster) formation and run simulations 
that reach $z=3.3$, and the run of \citet{2017MNRAS.465.3622R} reaches $z=0.5$. 
\citet{2018MNRAS.474.4232K} find that mergers can push gas to high density that
quickly forms clustered stars that end up tightly bound by the end of the 
simulation. A somewhat different approach couples semi-analytical models to DM-only
simulations \citep{2010ApJ...718.1266M,2014ApJ...796...10L,2018MNRAS.480.2343C,
2019MNRAS.486..331C,2019arXiv190505199C}, also done in the work by 
\citet{2019MNRAS.482.4528E}.

@@

\subsection{Age-estimates}
\label{sec:age}
The top panel of Figure~\ref{fig:MW-M31-age-Rgc} shows a histogram of the age-estimates
of the $55$ MW GCs in V13 and $88$ GCs in M31 for which age-estimates are available
in CR16. The mean age of the MW GCs in this data set is $11.9$~Gyr and the 
standard deviation is $0.8$~Gyr. Furthermore, only 1 of the 55 GC age-estimates is
below $10$~Gyr. The M31 GCS has a mean age of $11.0$~Gyr with a standard deviation 
of $2.2$~Gyr, and $24$ GCs have age-estimates below $10$~Gyr with a minimum age 
of $4.8$~Gyr. Based on these data, we find that the age distributions of the MW 
and M31 GCSs are not statistically consistent. The MW appears 
to host a GCS that is somewhat older than that of M31.
This is somewhat surprising given that M31 is generally considered to be earlier
type than the MW so that an older stellar population would naively be expected.
However, we do caution that both data sets are incomplete and the age 
measurements have large uncertainties (of $1-2$~Gyr). On the other hand, the 
magnitude of the uncertainty is insufficient to explain the low-age tail in M31.

@@

\section{Results}
\label{sec:results}
We define GC candidates in the Auriga simulations as all star particles older
than $10$~Gyr based on the age distribution of the MW GCS (top panel of 
Figure~\ref{fig:MW-M31-age-Rgc}), adopting the definition of a GC from
\citet{2010ARA&A..48..431P} and following the method of 
\citet{2017MNRAS.465.3622R}. However, we caution that his definition, although 
statistically consistent with the estimated ages of all GCs in the MW, is not 
consistent with the tail of younger clusters as stated in Section~\ref{sec:age},
and further discussed in Section~\ref{sec:age_min}.

@@

\begin{figure}
    \includegraphics[width=\columnwidth]{{FeH_obs_Au4-4_Au4-10_Au4-21-trim}.png}
    \caption{
        \emph{Top}: Metallicity distribution of the MW and M31 GCS. The solid 
        purple (magenta) line shows the GCSs of the MW (M31), and we 
        indicate the mean value with an arrow at the bottom of the x-axis. We
        overplot a double Gaussian for the MW GCS using purple dotted lines, taking
        the mean and dispersion from \citep{Harris2001}.
        \emph{Other panels}: three selected Auriga haloes as indicated in the 
        upper right corner. In each of these three panels we show the GG 
        candidates in beige, and use a beige dotted line for all star particles. 
        We split the GC candidates up into two subpopulations, those that have 
        formed {\it in~situ} (brown), and those that have been accreted (teal).
        The numbers in parenthesis show how many star particles are included 
        in the subset.
        \label{fig:Au4-10and21_FeH}
    }
\end{figure}
We select \mbox{Au4-10}\footnote{The nomenclature is `Au' for Auriga, followed by the
resolution level (4) and halo number (10, indicating which set of initial
conditions was used).}, \mbox{Au4-21}, and \mbox{Au4-4} to show the
distributions of three individual simulation runs that we consider representative
for specific behaviour, and to highlight that different runs give rise to 
different distributions. We show the age-selected GC candidates as well as the 
accreted and {\it in~situ} subsets. 
The top panel of Figure~\ref{fig:Au4-10and21_FeH} shows the MW 
and M31 GCSs, where we overplot a double Gaussian for the MW GCS by 
adopting literature values of the mean $\mu$ and standard deviation $\sigma$ of 
the metal-rich and metal-poor subpopulations \citep{Harris2001}.

@@

We show the mean metallicity and standard deviation of all $30$~Auriga L4 haloes
in Figure~\ref{fig:FeH_mu_sigma} to illustrate the individual behaviour for the 
full set of Auriga galaxies. The beige crosses are to be compared to the purple 
and magenta crosses, which show the mean value of all MW and M31 GCs, respectively. 
In addition, we show the metal-rich population of the MW GCS using a solid dot, 
and the metal-poor population with an open dot. We caution that
these literature values result from individual Gaussian fits to subsets of the 
observational data cut at [Fe/H]~$= -1$. We do not include corresponding data 
points calculated using such an artificial cut for the simulated star particles.
It is interesting that the mean metallicities of the {\it in~situ} GC candidates 
appear roughly consistent with that of the metal-rich population of the MW despite 
the lack of such a metallicity cut, although the simulations show larger dispersions. 
The latter could simply be caused by the hard separation of the MW data into two 
groups, which means the range is smaller and the resulting dispersion lower. With 
regard to the M31 GCS, so far no definitive consensus has been reached
in the literature concerning uni- bi- or trimodality in the [Fe/H] distribution, 
but CR16 argues that the data, after removal of younger objects due to improved 
age classification, hints at three populations separated at [Fe/H]~$=-0.4$ and 
[Fe/H]~$ = 1.5$. None of the Auriga subsets has a mean value that offers much 
hope of reconciling the simulated distribution with the lowest metallicity 
group in the M31 GCS. This is also true for the metal-poor population of the 
MW with a mean value of $-1.6$. The main result of this figure is 
that all Auriga L4 galaxies have metallicity distributions with (much) larger
mean values than observed for the MW and M31 GCS, and that we 
systematically find lower mean metallicities for the accreted GC candidates
than for those that have formed {\it in~situ}.

@@

The qualitative behaviour of the median distribution of galactocentric radii 
(Figure~\ref{fig:Rgc}) is unaffected by lowering the age cut other than including
more star particles (more mass) which leads to a higher normalisation and
a higher mass ratio with respect to the observed GC systems of the MW and M31.
The qualitative behaviour presented in Figure~\ref{fig:Au-FeHRgc-ratio} does not 
change when the age cut is lowered to 8 or 6 Gyr. 
GC candidates is not affected definition because our selection
function mostly includes star particles that have formed before roughly $z=2$.
Overall, reducing the age cut worsens the agreement between the observed GCSs 
and the candidate GCs in the Auriga galaxies.

>>>>

\section{Introduction}
Globular star clusters (GC)s are old, bright, and ubiquitous. Various properties of
GC systems (GCS) show correlations with their host galaxies. GCSs are believed to
retain information about the galactic (gas) conditions at their times of formation.
Thus they could offer a unique insight into the (chemodynamical) evolution of their 
parent galaxies, if the formation and evolution of GCs- and GC systems themselves 
is adequately understood. However, despite decades of research (e.g.~\citealt{
1991ARA&A..29..543H, Harris2001, 2004Natur.427...31W, 2006ARA&A..44..193B, 
2012A&ARv..20...50G, 2014CQGra..31x4006K, 2018RSPSA.47470616F}), consensus on an 
exhaustive picture of the formation of GCs is yet to be reached. Most galaxies host
a GCS with a bimodal colour distribution that may reflect a bimodal metallicity 
distribution which is argued to arise from two distinct formation channels 
\citep{2006ARA&A..44..193B}. GCs are typically referred to as `blue' or `red' \citep[e.g.][]{
1985ApJ...293..424Z,1999AJ....118.1526G,2001AJ....121.2974L} where the former class
is more metal-poor, more radially extended, and showing little to no rotation whereas
the latter is more metal-rich, radially concentrated, and rotating with the galaxy.

@@

Yet another hypothesis is that star cluster formation traces periods of high-intensity
star formation, which can be triggered by major gas-rich (spiral) galaxy mergers 
\citep{1987nngp.proc...18S, 1992ApJ...384...50A}. Such a scenario is naturally 
expected within the framework of hierarchical assembly and predicts formation 
of young clusters in interacting and merging galaxies, which has been observed 
and are found to show remarkable similarities with GCs in the Milky Way (MW) \citep[e.g.][]{
1995AJ....109..960W, 1996AJ....112..416H, 1999AJ....118..752Z, 1999AJ....118.1551W}.
Moreover, modelling efforts of this framework yield GC (sub)populations consistent 
with various observables \citep[e.g.][]{2010ApJ...718.1266M, 2018MNRAS.480.2343C}, 
and the recent numerical simulation of an isolated dwarf-dwarf merger executed at 
very high resolution (baryonic mass $m_b \sim$~4M\Sun; softening $\epsilon = 0.1$~pc) 
produces star clusters that could be GC progenitors \citep{
2019arXiv190509840L}.

@@

The (mass) resolution of cosmological zoom simulations has reached the mass range 
populated by GCs, and the gravitational force softening can be as low as several 
parsec. A number of groups can thus incorporate formation of (globular) star 
clusters into their high-resolution hydrodynamical simulations. \citet{2016ApJ...831..204R} 
ran parsec scale simulations of the high-redshift universe 
prior to reionization (the simulations stop at $z=9$), \citet{2017ApJ...834...69L} 
implement a new subgrid model for star (cluster) formation and run simulations 
that reach $z=3.3$, and the run of \citet{2017MNRAS.465.3622R} reaches $z=0.5$. 
\citet{2018MNRAS.474.4232K} find that mergers can push gas to high density that
quickly forms clustered stars that end up tightly bound by the end of the 
simulation. A somewhat different approach couples semi-analytical models to DM-only
simulations \citep{2010ApJ...718.1266M,2014ApJ...796...10L,2018MNRAS.480.2343C,
2019MNRAS.486..331C,2019arXiv190505199C}, also done in the work by 
\citet{2019MNRAS.482.4528E}.

@@

\subsection{Age-estimates}
\label{sec:age}
The top panel of Figure~\ref{fig:MW-M31-age-Rgc} shows a histogram of the age-estimates
of the $55$ MW GCs in V13 and $88$ GCs in M31 for which age-estimates are available
in CR16. The mean age of the MW GCs in this data set is $11.9$~Gyr and the 
standard deviation is $0.8$~Gyr. Furthermore, only 1 of the 55 GC age-estimates is
below $10$~Gyr. The M31 GCS has a mean age of $11.0$~Gyr with a standard deviation 
of $2.2$~Gyr, and $24$ GCs have age-estimates below $10$~Gyr with a minimum age 
of $4.8$~Gyr. Based on these data, we find that the age distributions of the MW 
and M31 GCSs are not statistically consistent. The MW appears 
to host a GCS that is somewhat older than that of M31.
This is somewhat surprising given that M31 is generally considered to be earlier
type than the MW so that an older stellar population would naively be expected.
On the other hand, the MW could be an outlier as it may have formed and assembled 
most of its mass earlier than galaxies with a similar mass \citep[e.g.][]{2014ApJ...781L..31S,
2015A&A...578A..87S,2018MNRAS.477.5072M}.
In addition, we caution that both data sets are incomplete and the age 
measurements have large uncertainties (of $1-2$~Gyr). On the other hand, the 
magnitude of the uncertainty is insufficient to explain the low-age tail in M31.

@@

\section{Results}
\label{sec:results}
We define GC candidates in the Auriga simulations as all star particles older
than $10$~Gyr based on the age distribution of the MW GCS (top panel of 
Figure~\ref{fig:MW-M31-age-Rgc}), adopting the definition of a GC from
\citet{2010ARA&A..48..431P} and following the method of 
\citet{2017MNRAS.465.3622R}. However, we caution that this definition, although 
statistically consistent with the estimated ages of all GCs in the MW, is not 
consistent with the tail of younger clusters as stated in Section~\ref{sec:age},
and further discussed in Section~\ref{sec:age_min}.

@@

\begin{figure}
    \includegraphics[width=\columnwidth]{{FeH_obs_Au4-4_Au4-10_Au4-21-trim}.png}
    \caption{
        \emph{Top}: Metallicity distribution of the MW and M31 GCS. The solid 
        purple (magenta) line shows the GCSs of the MW (M31), and we 
        indicate the mean value with an arrow at the bottom of the x-axis. We
        overplot a double Gaussian for the MW GCS using purple dotted lines, taking
        the mean and dispersion from \citep{Harris2001}.
        \emph{Other panels}: three selected Auriga haloes as indicated in the 
        upper right corner. In each of these three panels we show the GG 
        candidates in beige, and use a beige dotted line for all star particles. 
        We split the GC candidates up into two subpopulations, those that have 
        formed {\it in~situ} (brown), and those that have been accreted (teal).
        The numbers in parenthesis show how many star particles are included 
        in the subset.
        \label{fig:Au4-10and21_FeH}
    }
\end{figure}
We select \mbox{Au4-10}\footnote{The nomenclature is `Au' for Auriga, followed by the
resolution level 4 and halo number 10, indicating which set of initial
conditions was used.}, \mbox{Au4-21}, and \mbox{Au4-4} to show the
distributions of three individual simulation runs that we consider representative
for specific behaviour, and to highlight that different runs give rise to 
different distributions. We show the age-selected GC candidates as well as the 
accreted and {\it in~situ} subsets. 
The top panel of Figure~\ref{fig:Au4-10and21_FeH} shows the MW 
and M31 GCSs, where we overplot a double Gaussian for the MW GCS by 
adopting literature values of the mean $\mu$ and standard deviation $\sigma$ of 
the metal-rich and metal-poor subpopulations \citep{Harris2001}.

@@

We show the mean metallicity and standard deviation of all $30$~Auriga L4 haloes
in Figure~\ref{fig:FeH_mu_sigma} to illustrate the individual behaviour for the 
full set of Auriga galaxies. The beige crosses are to be compared to the purple 
and magenta crosses, which show the mean value of all MW and M31 GCs, respectively. 
In addition, we show the metal-rich population of the MW GCS using a solid dot, 
and the metal-poor population with an open dot. We caution that
these literature values result from individual Gaussian fits to subsets of the 
observational data cut at [Fe/H]~$= -1$. We do not include corresponding data 
points calculated using such an artificial cut for the simulated star particles.
It is interesting that the mean metallicities of the {\it in~situ} GC candidates 
appear roughly consistent with that of the metal-rich population of the MW despite 
the lack of such a metallicity cut, although the simulations show larger dispersions. 
The latter could simply be caused by the hard separation of the MW data into two 
groups, which means the range is smaller and the resulting dispersion lower. With 
regard to the M31 GCS, so far no definitive consensus has been reached
in the literature concerning uni- bi- or trimodality in the [Fe/H] distribution, 
but CR16 argues that the data, after removal of younger objects due to improved 
age classification, hints at three populations separated at [Fe/H]~$=-0.4$ and 
[Fe/H]~$ = -1.5$. None of the Auriga subsets has a mean value that offers much 
hope of reconciling the simulated distribution with the lowest metallicity 
group in the M31 GCS. This is also true for the metal-poor population of the 
MW with a mean value of $-1.6$. The main result of this figure is 
that all Auriga L4 galaxies have metallicity distributions with (much) larger
mean values than observed for the MW and M31 GCS, and that we 
systematically find lower mean metallicities for the accreted GC candidates
than for those that have formed {\it in~situ}.

@@

The qualitative behaviour of the median distribution of galactocentric radii 
(Figure~\ref{fig:Rgc}) is unaffected by lowering the age cut other than including
more star particles (more mass) which leads to a higher normalisation and
a higher mass ratio with respect to the observed GC systems of the MW and M31.
The qualitative behaviour presented in Figure~\ref{fig:Au-FeHRgc-ratio} does not 
change when the age cut is lowered to 8 or 6 Gyr. 
Overall, reducing the age cut worsens the agreement between the observed GCSs 
and the candidate GCs in the Auriga galaxies.
